sdmTMB: Fast, flexible, and user-friendly spatial and spatiotemporal generalized linear mixed-effects models

### 2021-11-02 doc/paper.Rmd
Update paper.Rmd

<!-- Paragraph on model selection here-->
Following glmmTMB, we allow AIC to be calculated for each sdmTMB model (Akaike XXXX). 
Because of well-documented biases of AIC with hierarchical or mixed-effects models (can cite Vaida and Blanchard or another), we also provide functionality to perform estimation via k-fold cross-validation. 
Our function `sdmTMB_cv()` allows for any number of folds or 'blocks', with assignments being user-specified or chosen randomly. 
The predictive log-density is calculated for each fold, and then summarized across folds for a total measure of the predictive accuracy of a model.

@@

The random fields are constructed as Gaussian Markov random fields (GMRF), where the random effects are structured as MVN constrained by a \mbox{Mat\'ern} covariance function.
<!-- EW: Because of space limits, I might suggest moving all this to the SI -->
The \mbox{Mat\'ern} defines the covariance $\Phi \left( s_j, s_k \right)$ between spatial locations $s_j$ and $s_k$ as $\Phi\left( s_j,s_k \right) = \tau^2/\Gamma(\nu)2^{\nu - 1}
(\kappa d_{jk})^\nu K_\nu \left( \kappa d_{jk} \right)$,
<!-- TODO: paraphrase this to avoid self plagiarizing the synopsis report! -->
where $\tau^2$ represents the spatial variance, $\Gamma$ represents the Gamma function, $K_\nu$ represents the Bessel function, $d_{jk}$ represents the Euclidean distance between locations $s_j$ and $s_k$, and $\kappa$ represents
a scaling parameter that is estimated (REF).
The parameter $\nu$ controls the smoothness of the covariance function.
We set $\nu = 1$, which lets us take advantage of the Stochastic Partial Differential Equation (SPDE) approximation to the GMRF to greatly increase computational efficiency [@lindgren2011].

@@

*EW: These are just placeholders -- not sure what order is best or what to highlight in very limited space*

@@

>>>>

<!-- Paragraph on model selection here-->
Following glmmTMB, we allow AIC to be calculated for each sdmTMB model [@akaike1974]. Because of the well documented biases of AIC with hierarchical or mixed-effects models [@liang2008] we also provide functionality to perform estimation via k-fold  cross-validation. Our wrapper function `sdmTMB_cv()` allows for any number of folds, with assignments being user-specified or chosen randomly. The predictive log-density is calculated for each fold, and then summarized across folds for a total measure of the predictive accuracy of a model.

@@

The random fields are constructed as Gaussian Markov random fields (GMRF), where the random effects are structured as MVN constrained by a \mbox{Mat\'ern} covariance function. Additional details are provided in the SI.

@@

@@

### Matern parameterization
<!-- EW: Because of space limits, I might suggest moving all this to the SI -->
The \mbox{Mat\'ern} defines the covariance $\Phi \left( s_j, s_k \right)$ between spatial locations $s_j$ and $s_k$ as $\Phi\left( s_j,s_k \right) = \tau^2/\Gamma(\nu)2^{\nu - 1}
(\kappa d_{jk})^\nu K_\nu \left( \kappa d_{jk} \right)$,
<!-- TODO: paraphrase this to avoid self plagiarizing the synopsis report! -->
where $\tau^2$ represents the spatial variance, $\Gamma$ represents the Gamma function, $K_\nu$ represents the Bessel function, $d_{jk}$ represents the Euclidean distance between locations $s_j$ and $s_k$, and $\kappa$ represents
a scaling parameter that is estimated (REF).
The parameter $\nu$ controls the smoothness of the covariance function.
We set $\nu = 1$, which lets us take advantage of the Stochastic Partial Differential Equation (SPDE) approximation to the GMRF to greatly increase computational efficiency [@lindgren2011].

### Junk to delete
*EW: These are just placeholders -- not sure what order is best or what to highlight in very limited space*
